Legacy Data Confounds Genomics Studies

### 2019-07-11 LatexManuscript/legacy.tex
figure reference

A recent publication by Mafessoni et al. also identified a batch effect in the 1kGP using a method that uses linkage disequilibrium rather than quality metrics to identify 19,196 suspicious variants with 67\% of them passing the 1kGP strict mask \citep{mafessoni2018turning} (Figure \ref{Mafessoni_MutSpect}).
They identify 17,917 variants significantly associated to abnormal LD patterns that are not associated to $Q$ \ref{Mafessoni_MutSpect}.
We find that 1,279 ( 3\% ) of the variants they identified are also in our list of suspicious variants and have correlated p-values to those identified using the GCAT method \ref{Mafessoni_MutSpect}.
We also find 23,111 $Q$-associated variants that are not associated to abnormal LD patterns. 
Interestingly, the variants identified my Mafessoni et al. are not enriched in the mutational spectra described above (Figure \ref{Mafessoni_MutSpect}).
These results indicate that there may a multitude of batch effects in the 1kGP that can only be identified using a suite of association tests.

>>>>

A recent publication by Mafessoni et al. also identified a batch effect in the 1kGP using a method that uses linkage disequilibrium rather than quality metrics to identify 19,196 suspicious variants with 67\% of them passing the 1kGP strict mask \citep{mafessoni2018turning} (Figure \ref{Mafessoni_MutSpect}A).
They identify 17,917 variants significantly associated to abnormal LD patterns that are not associated to $Q$.
We find that 1,279 ( 3\% ) of the variants they identified are also in our list of suspicious variants and have correlated p-values to those identified using the GCAT method.
We also find 23,111 $Q$-associated variants that are not associated to abnormal LD patterns. 
Interestingly, the variants identified my Mafessoni et al. are not enriched in the mutational spectra described above (Figure \ref{Mafessoni_MutSpect}B).
These results indicate that there may a multitude of batch effects in the 1kGP that can only be identified using a suite of association tests.
